Constructing Sailing Match Race Schedules: Round Robin Pairing Lists

### 2015-04-07 paper.tex
Formatting changes only

\section{Introduction}
%\coffee{1}
This work describes a round robin competition format that arises from sailing, known as match racing. Competitors, i.e. {\em skippers}, compete against each other in a series of matches taking place in rounds, known as {\em flights}.  A match is composed of two skippers, with each skipper in a boat on his own. Skippers in a match set off together, one on the port side, the other starboard, and first home wins that match. Criteria published by the Interational Sailing Federation (ISAF) dictate what makes a legal match race schedule and what should be optimised. This is a rich set of cirteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This is a daunting task. A close inspection of these schedules reveals that most are illegal, violating many of the match race scheduling criteria, many schedules are missing and some that are legal are far from optimal.

@@

\section{Problem Definition: Round Robin Pairing Lists}\label{sec:definition}
The guidelines for running match racing events~\cite{isaf} places a number of criteria on how the competing skippers are scheduled into flights and matches, known as pairing lists. Several of these criteria are applicable when the number of skippers exceeds the number of available boats, in which case skippers have to change in and out of boats, and are allowed time in the schedule to check and fine-tune the boat they change into. Typically, matches set off at 5 minute intervals with each match taking about 20 minutes. Therefore, if we have 10 skippers and 10 boats we have 45 matches. This results in 9 flights each of 5 matches.  Each flight takes about 50 minutes, consequently eight or nine flights a day is a reasonable achievement for most events~\cite{isaf}. Consequently, the number of boats and skippers involved is typically small.

@@

\begin{framed}
Principal Criteria in Order of Priority:
\begin{enumerate}
\item Each skipper sails against each other skipper once.
\item *When skippers have an even number of matches, they have the same number of port and starboard assignments.
\item *When skippers have an odd number of matches, the first half of the skippers
    will have one more starboard assignment.
%\item When there is an even number of flights, each skipper has the same number
%    of port and starboard assignments.
%\item When there is an odd number of flights, the first half of the skippers
%    will have one more starboard assignment.
\item No skipper in the last match of a flight should be in the first match of
    the next flight.
\item No skipper should have more than two consecutive port or starboard
    assignments.
\item Each skipper should be assigned to match 1, match 2, etc. in a flight as
    equally as possible.
\item In flights with five or more matches, no skipper should be in the
    next-to-last match in a flight and then in the first match of the next
    flight.
\item If possible, a skipper should be starboard when meeting the nearest
    lowest ranked skipper (i.e.\ \#1 will be starboard against \#2, \#3 will be
    starboard against \#4).
\item Close-ranked skippers meet in the last flight.
\item Minimise the number of boat changes.
\item Skippers in the last match of a flight do not change boats.
\item Skippers in new boats do not sail in the first match of the next flight.
\item Skippers have a reasonable sequence of matches and blanks.
\end{enumerate}
Note that criteria 10, 11 and 12 only apply when there are fewer boats than
skippers; 11 and 12 override 6 when changes are required.
\end{framed}

@@

\begin{figure}[tb]
\centering
\includegraphics[height=10.2cm,width=13.2cm]{oriented.pdf}
\caption{A pictorial representation of the orientation process, stage 4. Top
right is the actual oriented schedule. Below that is the dfa to satisfy
criteria 5. On the left is the zero/one variables, a row representing a
skippers sequence of competitors. An edge between variables corresponds to a
matched pair, that must take different values. Bottom right is a sketch by third author.}
\label{oriented}
\end{figure}

\section{Sample Schedules}\label{sec:samples}
We present three pairs of schedules. The first pair is for 8 skippers and 6 boats, one published by ISAF, the other generated by us. Next, we present two new schedules, not appearing in the ISAF manual, for 9 skippers and 6 boats and 9 skippers with 8 boats. The last pair is for 10 skippers and 6 boats, again ISAF published and generated by us. For each of our schedules, we used 1 day cpu time for the first stage (minimising boat changes) then a further 2 days to minimise imbalance. Stages 3 and 4 (renaming and orientation) completed in less than a second.

@@

\noindent
\subsection{8 skippers and 6 boats}
We first analyse the ISAF schedule, Table \ref{08-06} on the left. Criteria 4 (skippers in last match in a flight cannot be first in next flight) is violated on 3 occasions (skipper 1 in flight 3, skipper 7 in flight 4, skipper 0 in flight 7). Criteria 5 (no more than two consecutive port or starboard assignments) is violated for skipper 6 (flights 1 to 3) and 7 (flights 7 to 9). Criteria 6 (imbalance) is violated for skippers 6 and 8. Criteria 12  (new boats do not sail first) is violated in flight 9 for skipper 5. Finally, the schedule has 8 boat changes. Our schedule, Table \ref{08-06} on the right, respects all criteria and has 6 boat changes and a maximum imbalance of 1.

\subsection{9 skippers}
Two new schedules are presented in Table \ref{09}, on the left for 6 boats and on the right 8 boats. Neither of these schedules appear in the ISAF Umpires' Manual. Both schedules respect all criteria. For 6 boats there are 8 boat changes (no skipper making more than 1 boat change) and a maximum imbalance of 2. For 8 boats there are again 8 boat changes, no skipper with more than 1 boat change, and perfect balance (each skipper 4 times on starboard side and 4 on port side).

@@

\subsection{10 skippers and 6 boats}
Table \ref{10-06} shows the ISAF schedule on the right for 10 skippers and 6 boats and on the right, our schedule.  The ISAF schedule violates criteria 5 (no more than 2 consequitive port or starboard assignments) 10 times (twice for skipper 0, 1 and 6 and once for skippers 3, 7, 8 and 9).  Criteria 12 (new boats do not sail first) is violated 7 times (in flight 3 for skippers 6 and 2, flight 5 for skipper 8, flight 6 for skipper 7, flight 8 for skipper 1, flight 9 for skipper 4 and flight 12 for skipper 4 again). There are 22 boat changes, with the minimum changes for a skipper being 1 and the maximum 3. Our schedule, right of Table \ref{10-06} satisfies all criteria, has 12 boat changes with the minimum changes for a skipper being 0 and the maximum 2, and a maximum imbalance of 1.

@@

\subsection{Discussion}
It was anticipated that this problem would be easy to model. This naivety was due to "round robin", leading us to believe that we could use some of the models proposed in \cite{trick2002}, \cite{trick2008} and \cite{henz2004}. This naivety slowly dissolved as we addressed more and more of the ISAF criteria. A number of models we produced were scrapped due to misunderstanding about the actual problem, i.e. a communication problem between the authors. Eventually this was resolved by presenting the majority of the criteria as a deterministic finite automaton. This became our Rosetta Stone, with the surprising benefit that it not only improved the teams communications, it was also a constraint!

@@

\section{Conclusion}\label{sec:conclusions}
We have produce new and better match race schedules. These schedules can be used by anyone that competes under the criteria published by ISAF. Our schedules can be downloaded as blank schedules and then populated with the names of the actual competing skippers.

@@

%\section*{Acknowledgments}
%We would like to thank our parents, God, and all his angels.
%\coffee{2}

>>>>

\section{Introduction}

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e. {\em skippers}, compete against each other in a series of matches taking
place in rounds, known as {\em flights}.  A match is composed of two skippers, with each skipper in
a boat on his own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the Interational Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
cirteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing and some that are legal are
far from optimal.

@@

\section{Problem Definition: Round Robin Pairing Lists}\label{sec:definition}

The guidelines for running match racing events~\cite{isaf} places a number of criteria on how the
competing skippers are scheduled into flights and matches, known as pairing lists. Several of these
criteria are applicable when the number of skippers exceeds the number of available boats, in which
case skippers have to change in and out of boats, and are allowed time in the schedule to check and
fine-tune the boat they change into. Typically, matches set off at 5 minute intervals with each
match taking about 20 minutes. Therefore, if we have 10 skippers and 10 boats we have 45 matches.
This results in 9 flights each of 5 matches.  Each flight takes about 50 minutes, consequently eight
or nine flights a day is a reasonable achievement for most events~\cite{isaf}. Consequently, the
number of boats and skippers involved is typically small.

@@

\begin{framed}
Principal Criteria in Order of Priority:
\begin{enumerate}
\item Each skipper sails against each other skipper once.
\item *When skippers have an even number of matches, they have the same number of port and starboard
    assignments.
\item *When skippers have an odd number of matches, the first half of the skippers will have one
    more starboard assignment.
\item No skipper in the last match of a flight should be in the first match of the next flight.
\item No skipper should have more than two consecutive port or starboard assignments.
\item Each skipper should be assigned to match 1, match 2, etc. in a flight as equally as possible.
\item In flights with five or more matches, no skipper should be in the next-to-last match in a
    flight and then in the first match of the next flight.
\item If possible, a skipper should be starboard when meeting the nearest lowest ranked skipper
    (i.e.\ \#1 will be starboard against \#2, \#3 will be starboard against \#4).
\item Close-ranked skippers meet in the last flight.
\item Minimise the number of boat changes.
\item Skippers in the last match of a flight do not change boats.
\item Skippers in new boats do not sail in the first match of the next flight.
\item Skippers have a reasonable sequence of matches and blanks.
\end{enumerate}
Note that criteria 10, 11 and 12 only apply when there are fewer boats than skippers; 11 and 12
override 6 when changes are required.
\end{framed}

@@

\begin{figure}[tb]
\centering
\includegraphics[height=10.2cm,width=13.2cm]{oriented.pdf}
\caption{A pictorial representation of the orientation process, stage 4. Top right is the actual
oriented schedule. Below that is the dfa to satisfy criteria 5. On the left is the zero/one
variables, a row representing a skippers sequence of competitors. An edge between variables
corresponds to a matched pair, that must take different values. Bottom right is a sketch by third
author.}\label{oriented}
\end{figure}

\section{Sample Schedules}\label{sec:samples}

We present three pairs of schedules. The first pair is for 8 skippers and 6 boats, one published by
ISAF, the other generated by us. Next, we present two new schedules, not appearing in the ISAF
manual, for 9 skippers and 6 boats and 9 skippers with 8 boats. The last pair is for 10 skippers and
6 boats, again ISAF published and generated by us. For each of our schedules, we used 1 day cpu time
for the first stage (minimising boat changes) then a further 2 days to minimise imbalance. Stages 3
and 4 (renaming and orientation) completed in less than a second.

@@

\noindent
\subsection{8 skippers and 6 boats}

We first analyse the ISAF schedule, Table \ref{08-06} on the
left. Criteria 4 (skippers in last match in a flight cannot be first in next flight) is violated on
3 occasions (skipper 1 in flight 3, skipper 7 in flight 4, skipper 0 in flight 7). Criteria 5 (no
more than two consecutive port or starboard assignments) is violated for skipper 6 (flights 1 to 3)
and 7 (flights 7 to 9). Criteria 6 (imbalance) is violated for skippers 6 and 8. Criteria 12  (new
boats do not sail first) is violated in flight 9 for skipper 5. Finally, the schedule has 8 boat
changes. Our schedule, Table \ref{08-06} on the right, respects all criteria and has 6 boat changes
and a maximum imbalance of 1.

\subsection{9 skippers}

Two new schedules are presented in Table \ref{09}, on the left for 6 boats
and on the right 8 boats. Neither of these schedules appear in the ISAF Umpires' Manual. Both
schedules respect all criteria. For 6 boats there are 8 boat changes (no skipper making more than 1
boat change) and a maximum imbalance of 2. For 8 boats there are again 8 boat changes, no skipper
with more than 1 boat change, and perfect balance (each skipper 4 times on starboard side and 4 on
port side).

@@

\subsection{10 skippers and 6 boats}

Table \ref{10-06} shows the ISAF schedule on the right for 10 skippers and 6 boats and on the right,
our schedule.  The ISAF schedule violates criteria 5 (no more than 2 consequitive port or starboard
assignments) 10 times (twice for skipper 0, 1 and 6 and once for skippers 3, 7, 8 and 9).  Criteria
12 (new boats do not sail first) is violated 7 times (in flight 3 for skippers 6 and 2, flight 5 for
skipper 8, flight 6 for skipper 7, flight 8 for skipper 1, flight 9 for skipper 4 and flight 12 for
skipper 4 again). There are 22 boat changes, with the minimum changes for a skipper being 1 and the
maximum 3. Our schedule, right of Table \ref{10-06} satisfies all criteria, has 12 boat changes with
the minimum changes for a skipper being 0 and the maximum 2, and a maximum imbalance of 1.

@@

\subsection{Discussion}

It was anticipated that this problem would be easy to model. This naivety was due to "round robin",
leading us to believe that we could use some of the models proposed in \cite{trick2002},
\cite{trick2008} and \cite{henz2004}. This naivety slowly dissolved as we addressed more and more of
the ISAF criteria. A number of models we produced were scrapped due to misunderstanding about the
actual problem, i.e. a communication problem between the authors. Eventually this was resolved by
presenting the majority of the criteria as a deterministic finite automaton. This became our Rosetta
Stone, with the surprising benefit that it not only improved the teams communications, it was also a
constraint!

@@

\section{Conclusion}\label{sec:conclusions}

We have produce new and better match race schedules. These schedules can be used by anyone that
competes under the criteria published by ISAF. Our schedules can be downloaded as blank schedules
and then populated with the names of the actual competing skippers.

@@
